Clique and Constraint Models for Maximum Common (Connected) Subgraph Problems

### 2016-04-20 paper/paper.tex
See if we can fit the association graph figure in

An alternative approach to MCS is to reduce the problem to finding a maximum clique in an
association graph \cite{LeviG,bala86,dura99,DBLP:journals/jcamd/RaymondW02a}.
An association graph (or compatibility graph, or weak modular product) of two graphs $G$ and $H$ is
an undirected graph $G \modularproduct H$ with vertex set $\operatorname{V}(G \modularproduct H) =
\operatorname{V}(G) \times \operatorname{V}(H)$---to avoid confusing vertices of $G \modularproduct
H$ with vertices of the two original graphs, we call vertices of $G \modularproduct H$
\emph{matching nodes}, as each vertex $(u, u')$ of $G \modularproduct H$ denotes the matching of $u$
with $u'$. The edges of $G \modularproduct H$ connect matching nodes which denote compatible
assignments, so two matching nodes $(u, u')$ and  $(v, v')$ are adjacent if $u \neq v$ and $u' \neq
v'$, and if they preserve both edges and non-edges, so $(u, v) \in \operatorname{E}(G) \Leftrightarrow
(u', v') \in \operatorname{E}(H)$.

@@

?? Can we figure out why? Is it time-to-find or time-to-prove? Assume that this will take up a few
sentences of space.

>>>>

An alternative approach to MCS is to reduce the problem to finding a maximum clique in an
association graph \cite{LeviG,bala86,dura99,DBLP:journals/jcamd/RaymondW02a}.
An association graph (or compatibility graph, or weak modular product) of two graphs $G$ and $H$ is
an undirected graph $G \modularproduct H$ with vertex set $\operatorname{V}(G \modularproduct H) =
\operatorname{V}(G) \times \operatorname{V}(H)$---to avoid confusing vertices of $G \modularproduct
H$ with vertices of the two original graphs, we call vertices of $G \modularproduct H$
\emph{matching nodes}, as each vertex $(u, u')$ of $G \modularproduct H$ denotes the matching of $u$
with $u'$. The edges of $G \modularproduct H$ connect matching nodes which denote compatible
assignments, so two matching nodes $(u, u')$ and  $(v, v')$ are adjacent if $u \neq v$ and $u' \neq
v'$, and if they preserve both edges and non-edges, so $(u, v) \in \operatorname{E}(G) \Leftrightarrow
(u', v') \in \operatorname{E}(H)$. We illustrate this in \cref{figure:association}.

\begin{figure}[tb]
    \centering
    \begin{tikzpicture}[scale=0.33]%{{{
        \begin{scope}
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Ma) at (2, 10) {\phantom{0}};
            \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (Mb) at (0, 8) {\phantom{0}};
            \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (Mc) at (2, 8) {\phantom{0}};
            \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (Md) at (4, 8) {\phantom{0}};

            \node [right = 0 of Ma, font=\small] { \vphantom{0}a };
            \node [below = 0 of Mb, font=\small] { \vphantom{0}b };
            \node [below = 0 of Mc, font=\small] { \vphantom{0}c };
            \node [below = 0 of Md, font=\small] { \vphantom{0}d };

            \draw [very thick] (Mb) -- (Mc);
            \draw [very thick] (Mc) -- (Md);
            \draw [color=uofgsandstone!40] (Ma) -- (Mc);
            \draw [color=uofgsandstone!40] (Ma) to [out=150, in=80, looseness=8] (Ma);
            \draw [very thick] (Mb) to [out=150, in=80, looseness=8] (Mb);
        \end{scope}

        \begin{scope}
            \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (M1) at (1, 2) {\phantom{0}};
            \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (M2) at (3, 2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (M3) at (1, 0) {\phantom{0}};
            \node[draw, circle, fill=chromayellow, inner sep=0.5pt, font=\normalsize] (M4) at (3, 0) {\phantom{0}};

            \node [left = 0 of M1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of M2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of M3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of M4, font=\small] { \vphantom{0}4 };

            \draw [very thick] (M1) -- (M2);
            \draw [very thick] (M2) -- (M4);
            \draw [color=uofgsandstone!40] (M3) -- (M4);
            \draw [color=uofgsandstone!40] (M1) -- (M3);

            \draw [color=uofgsandstone!40] (M3) to [out=280, in=210, looseness=8] (M3);
            \draw [very thick] (M1) to [out=150, in=80, looseness=8] (M1);
        \end{scope}

        \begin{scope}[xshift=15cm]
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Ma1) at (2.6, 9) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Ma2) at (4.2, 9) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Ma3) at (5.8, 9) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Ma4) at (7.4, 9) {\phantom{0}};
            \node [above = 0 of Ma1, font=\small] { \vphantom{0}1 };
            \node [above = 0 of Ma2, font=\small] { \vphantom{0}2 };
            \node [above = 0 of Ma3, font=\small] { \vphantom{0}3 };
            \node [above = 0 of Ma4, font=\small] { \vphantom{0}4 };

            \draw [decorate, decoration={brace, raise=0.5cm}, very thick] (Ma1.north west) -- (Ma4.north east)
            node [midway, above=0.7cm] { $a \mapsto$ };

            \node[draw, circle, fill=chromagreen, inner sep=0.5pt, font=\normalsize] (Mb1) at (1, 2.6) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mb2) at (1, 4.2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mb3) at (1, 5.8) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mb4) at (1, 7.4) {\phantom{0}};
            \node [left = 0 of Mb1, font=\small] { \vphantom{0}1 };
            \node [left = 0 of Mb2, font=\small] { \vphantom{0}2 };
            \node [left = 0 of Mb3, font=\small] { \vphantom{0}3 };
            \node [left = 0 of Mb4, font=\small] { \vphantom{0}4 };

            \draw [decorate, decoration={brace, raise=0.5cm}, very thick] (Mb1.south west) -- (Mb4.north west)
            node [midway, left=0.7cm] { $b \mapsto$ };

            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc1) at (2.6, 1) {\phantom{0}};
            \node[draw, circle, fill=chromagreen, inner sep=0.5pt, font=\normalsize] (Mc2) at (4.2, 1) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc3) at (5.8, 1) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Mc4) at (7.4, 1) {\phantom{0}};
            \node [below = 0 of Mc1, font=\small] { \vphantom{0}1 };
            \node [below = 0 of Mc2, font=\small] { \vphantom{0}2 };
            \node [below = 0 of Mc3, font=\small] { \vphantom{0}3 };
            \node [below = 0 of Mc4, font=\small] { \vphantom{0}4 };

            \draw [decorate, decoration={brace, raise=0.5cm}, very thick] (Mc4.south east) -- (Mc1.south west)
            node [midway, below=0.7cm] { $c \mapsto$ };

            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Md1) at (9, 2.6) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Md2) at (9, 4.2) {\phantom{0}};
            \node[draw, circle, fill=white, draw=uofgsandstone!40, inner sep=0.5pt, font=\normalsize] (Md3) at (9, 5.8) {\phantom{0}};
            \node[draw, circle, fill=chromagreen, inner sep=0.5pt, font=\normalsize] (Md4) at (9, 7.4) {\phantom{0}};
            \node [right = 0 of Md1, font=\small] { \vphantom{0}1 };
            \node [right = 0 of Md2, font=\small] { \vphantom{0}2 };
            \node [right = 0 of Md3, font=\small] { \vphantom{0}3 };
            \node [right = 0 of Md4, font=\small] { \vphantom{0}4 };

            \draw [decorate, decoration={brace, raise=0.5cm}, very thick] (Md4.north east) -- (Md1.south east)
            node [midway, right=0.7cm] { $d \mapsto$ };

            \begin{scope}[on background layer]
                \draw [draw=uofgsandstone!40] ($(Ma1.south)!0.5!(Ma1)$) -- ($(Md4.west)!0.5!(Md4)$);
                \draw [draw=uofgsandstone!40] ($(Ma1.south)!0.5!(Ma1)$) -- ($(Mc2.north)!0.5!(Mc2)$);
                \draw [draw=uofgsandstone!40] ($(Ma3.south)!0.5!(Ma3)$) -- ($(Mc4.north)!0.5!(Mc4)$);
                \draw [draw=uofgsandstone!40] ($(Mb3.east)!0.5!(Mb3)$) -- ($(Mc4.north)!0.5!(Mc4)$);
                \draw [draw=uofgsandstone!40] ($(Md2.west)!0.5!(Md2)$) -- ($(Mc4.north)!0.5!(Mc4)$);
                \draw [draw=uofgsandstone!40] ($(Md2.west)!0.5!(Md2)$) -- ($(Ma3.south)!0.5!(Ma3)$);
                \draw [draw=uofgsandstone!40] ($(Md2.west)!0.5!(Md2)$) -- ($(Mb3.east)!0.5!(Mb3)$);
                \draw [very thick] ($(Mb1.east)!0.5!(Mb1)$) -- ($(Mc2.north)!0.5!(Mc2)$);
                \draw [very thick] ($(Mb1.east)!0.5!(Mb1)$) -- ($(Md4.west)!0.5!(Md4)$);
                \draw [very thick] ($(Mc2.north)!0.5!(Mc2)$) -- ($(Md4.west)!0.5!(Md4)$);
        \end{scope}
        \end{scope}

    \end{tikzpicture}

    \caption{A maximum common induced subgraph between the two graphs on the left has three
        vertices---one solution is highlighted. On the right, the association graph encoding: the
        highlighted clique of size three shows the same solution. In practice the isolated vertices
        (representing assignments between incompatible vertices) would be omitted.}\label{figure:association}
\end{figure}

@@
